---
title: "Data Analaysis with the Arctic shorebird csv files"
format: html
---


```{r, message=FALSE}
library(tidyverse)
library(lubridate)
library(RColorBrewer)
```


## Define path and filenames

```{r}
path_raw <- "data/raw"

sites_csv <- "site.csv"

snowcover_csv <- "snow_cover.csv"

nests_csv <- "ASDN_Bird_nests.csv"

eggs_csv <- "ASDN_Bird_eggs.csv"

species_csv <- "species.csv"

lemmings_csv <- "ASDN_Daily_pred_lemm.csv"
```


## Question 1: Does the snow amount & timing impact the number of eggs? Is it species dependant?

### Summarize the nests table

Read the nest data in

```{r}
nests_data <- read_csv(file.path(path_raw, nests_csv))

glimpse(nests_data)
```

Yeah! `Date_found` is already in date format! We can thus directly aggregate the data by months:


```{r}
nests_month <- nests_data %>% 
  mutate(month = month(Date_found)) %>% 
  group_by(Species, Site, Year, month) %>% 
  summarize(n_nest = n())
```

### Summarize the snow cover table

Read the snow cover data in

```{r}
snowcover_data <- read_csv(file.path(path_raw, snowcover_csv))

glimpse(snowcover_data)
```

The Date is stored as a character we thus first we need to transform the date in date format before computing the monthly aggregate:

```{r}
snowcover_data_fixed <- snowcover_data %>% 
  mutate(Date2 = as_date(Date)) # Warning 72 failed to parse
# if failed, generally is converted to NA
```

We have 72 errors. Let's have a look at the date which failed:

```{r}
snowcover_data_fixed %>% 
  filter(is.na(Date2)) %>% 
  View()
```


It is because 2 days were entered as one value `8&9 june 06`, it is not in the expected format

```{r}
# fix the problems in date column and then rerun
snowcover_data_fixed <- snowcover_data_fixed %>% 
 mutate(Date = case_when(
   Date == "8&9 june 06" ~ "09-June-06",
   .default = Date
 )) %>% 
 mutate(Date2 = as_date(Date))
```

Mmm the Dates are not what we were expecting... It generally means there are different date types. Let's start by fixing the problematic dates before transforming it to the date format.

```{r}
snowcover_data_fixed <- snowcover_data_fixed %>% 
 mutate(Date = case_when(
   Date == "8&9 june 06" ~ "09-June-06",
   .default = Date
 )) %>% 
 mutate(Date2 = as_date(Date, format = "%d-%b-%y"))
```


::: {.callout-note}
## Fun fact
You might be wondering why I started with `as_date()` instead of directly using `dmy()`. Well turns out that while preparing this exercise I discovered a bug in this `lubridate` function!! See <https://github.com/tidyverse/lubridate/issues/1161>
:::


Now let's compute the monthly statistics
- if snow cover is larger than 10%
- group by site, year, month
- cumulative snow days (count)
- Average snow_cover (%)

```{r}
# adding the snow days with a threshold at 10%
# cumulative snow days and mean snow cover
snowcover_month <- snowcover_data_fixed %>% 
  mutate(month = month(Date2)) %>%
  mutate(snow_days = ifelse(Snow_cover > 10, 1, 0)) %>% 
  group_by(Site, Year, month) %>% 
  summarize(cum_snow_days = sum(snow_days, na.rm = TRUE),
            avg_snow_cover = mean(Snow_cover, na.rm = TRUE))
```

 
### Join the snow cover and nest tables

```{r}
nests_snow_month <- nests_month %>% 
  inner_join(snowcover_month, join_by(Site, month, Year))
```

**joins are one of the most expensive operations in R, always try to aggregate data BEFORE joining**

Well in the end we did not have much information by months (mainly 2 months per year focused on June and July), so let's aggregate by year instead:

```{r}
nests_snow_year <- nests_snow_month %>% 
  group_by(Species, Site, Year) %>% 
  summarize(n_nest = sum(n_nest),
            cum_snow_days = sum(cum_snow_days, na.rm = TRUE),
            avg_snow_cover = mean(avg_snow_cover, na.rm = TRUE))
```

### Plots

```{r}
ggplot(nests_snow_year, aes(x = cum_snow_days, y = n_nest, color = Species)) +
  geom_point() +
  scale_color_brewer(palette = "Set3") +
  theme_bw()
```

```{r}
# load species data so we know what the species code means
species_data <- read_csv(file.path(path_raw, species_csv))

nests_snow_year <- nests_snow_year %>% 
  left_join(species_data, join_by(Species==Code))
  
```

What do you think is the answer to our question? 


## Question 2: Which shorebird species makes the most eggs?

### Import the data

```{r}
sites_data <- read_csv(file.path(path_raw, sites_csv))

eggs_data <- read_csv(file.path(path_raw, eggs_csv))

species_data <- read_csv(file.path(path_raw, species_csv))
```

Let's compute the number of eggs per nest. The column `Egg-num` seems to be a good candidate to estimate the number of eggs per nest:

```{r}
eggs_sum <- 
```

Join with the nest data:

```{r}

```

Group by species:

```{r}

```

OK we might want to have more explicit names for the species!, Luckily we have the species table to help with that:

```{r}

```

Some cleaning of the data frame to put the average at the end and remove the Relevance column:

```{r}

```

